### 2018-07-17 Manuscripts/cog-psych/revision/cogpsych.tex
Added representativeness to manuscript. Updated figures and results. Discussion in various places and some reviewer-related responses still need work.

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

@@

% set up notes-- different backgrounds for Nolan and Joe!
\newcommand\nbcnote[1]{\todo[inline, backgroundcolor = yellow]{\textbf{NBC}: #1}}
\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}

@@

\vfill
Joseph L. Austerweil\textsuperscript{1}
Nolan Conaway\textsuperscript{1}, 
Shi X. Liew\textsuperscript{1},
and Kenneth J. Kurtz\textsuperscript{2}
\\[\baselineskip]
\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
\\[1in]

@@

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section*{Abstract}
\jlanote{need to update}
The ability to creatively generate new concepts and ideas is among the most fascinating aspects of human cognition, but we do not have a strong understanding of the cognitive processes and representations underlying concept generation. In this paper, we study the generation of new categories using the computational and behavioral toolkit of traditional artificial category learning. Previous work in this domain has focused on how the statistical structure of known categories generalizes to generated categories, overlooking the degree of contrast between the known and generated categories as a factor. We report two experiments demonstrating that contrast between what is known and what is created is of fundamental importance for categorization and the creative process. We propose a novel, exemplar-based approach explaining our results, and compare the model's performance to two key alternatives. Our experiments and simulations demonstrate specifically how category contrast influences creative generation, and how it can interact with other constraints to produce different types of creations. Our work also serves as a concrete example of how well-established and highly controlled experimental frameworks (such as those in the field of category learning) can be used to make progress in an intriguing yet understudied domain.

@@

To successfully generate something novel, what is generated must be different from what is already known. This fundamental constraint, ``being different'', or contrasting from other categories in the relevant domain, is the focus of our work. Although implicitly assumed in some work, this constraint has been overlooked in previous research: To our knowledge, there has not been any systematic investigation addressing how generated categories {\em differ} from what is already known. Although the idea of category contrast is ubiquitous throughout the categorization literature, and extends to a variety of other fields \citep[e.g., color;][]{regier2007}, the idea that a new category should be ``different'' is vague, as there are many ways it could be different from a previously observed category.

We examine two different definitions of how categories should differ from one another: exemplar contrast or representative of the alternative category. To formalize the first technique, we build on the largely successful exemplar modeling framework that \citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so, we propose a novel exemplar model of category generation, {\em Producing Alike and Contrasting Knowledge using Exemplar Representations} (PACKER), formalizing how new categories should differ from previous categories. It embodies the exemplar contrast principle by incorporating a factor that repulses members of opposite categories from one another. To foreshadow, we find that this notion of contrast predicts human category learning performance, above and beyond other category learning models in a novel behavioral experiment. This demonstrates that contrast plays a central role in other important categorization tasks, which is not specific to category generation.

The second hypothesis for how contrast might affect categorization is contrast as {\em representativeness}. The representativeness heuristic \citep{kahneman73} states that people make judgments regarding an outcome based on how representative it is of the evidence given in the current context. The heuristic is a powerful theoretical construct that has been used to capture a range of complex patterns of human judgments \citep{kahneman73,tversky74,tversky83}, especially those that deviate from normative theory (as given by a straightforward application of Bayes' rule). For example, the coinflip sequence TTHTH is perceived to be more random (due to it being representative of a coinflip sequence) than TTTTT (despite both having the same probability of having been generated from a fair coin). Although there is a healthy debate as to whether perceived randomness {\em really} deviates from normative theory \citep{griffiths18,hahn09}, the representativeness heuristic remains one of the main explanations of human judgments \citep{reimers18}. To formalize the hypothesis that people expect a new category to be representative of the opposite of the current category, or contrast as {\em representativeness}, we use the Bayesian formulation of representativeness \citep{tenenbaum01}, which has been used successfully to capture human performance in color categorization \citep{abbott16}, image category learning \citep{abbott11}, and language grammar learning \citep{rafferty10}. We do so by extending the state-of-the-art category generation model \citep{jern2013probabilistic} and find that  human category generation is best capture as generating representative, rather than probable, samples.

@@

@@

\jlanote{check to make sure the next paragraph has been updated enough}

Although it is evident that people are {\em capable} of creating new concepts and categories, it is not entirely clear how new concepts are systematically made different from what is already known. The hierarchical Bayesian model developed by \cite{jern2013probabilistic} assumes that differences between observed and generated categories are only due to random variation. The model assumes that generated categories are sampled from the same underlying domain distribution as observed categories, and will thus share a common distributional structure. The model does not make predictions about the {\em location} of the category within the domain (the perceptual instantiation of category members). Under a strict interpretation of their model, given knowledge of a single category within the domain, the most probable new category to be generated is located in {\em exactly} the same location and possesses an identical distributional structure. This is not an issue with their model specifically, but with a broader class of standard hierarchical Bayesian models \citep[e.g., ][]{griffithsscn07,kemp07}. Many of these models assume that at some point of the latent generative process the same underlying distribution generates all of the categories and thus, any differences between categories are due to {\em noise} and should not be {\em systematic}. If the goal of creating a new category given others in a domain is generating from the posterior predictive distribution then, the best a standard hierarchical Bayesian model without a notion of contrast built into the model can do at capturing contrast is to assume that the new category is placed uniformly at random over stimulus space. But, this defeats the purpose of a hierarchy as it is ignored when determining a new category location!\footnote{It is plausible that some hierarchical Bayesian model could be created that generates categories different from each other. However, this model would not be a standard application or extension of most pre-existing hierarchical Bayesian models. The generative process would need to include a component that promotes contrast.} Note that this does not mean location information and contrast cannot be captured using hierarchical Bayesian models. In fact, our model of contrast as representativeness is a hierarchical Bayesian model, but with a different goal than predicting the next category.

@@

\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}

@@

@@

PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership: $f(x_j) = \gamma$ if $x_j$ is a member of the target category, and $f(x_j) = \gamma - 1$ if $x_j$ is a member of a contrast category. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$) controlling the trade-off between within- and between-category similarity. For example, when $\gamma = 0.5$, $f(x_j) = 0.5$ for members of the target category and $f(x_j) = -0.5$ for members of other categories; thus, the model is likely to generate items that are similar to members of the target category but are not similar to members of other categories. In this way, $\gamma = 1$ produces exclusive consideration of target-category members, and $\gamma = 0$ produces exclusive consideration of opposite-category members. The $\gamma$ parameter thus specifies a wide breadth of possible approaches; by fitting it to a dataset, one can describe the relative roles of between-category contrast and within-category similarity in generation. See Figure \ref{fig:packer-examples} for an illustration of how $\gamma$ controls the relative influence of within category similarity and contrast to other categories when generating a new exemplar.

\begin{figure}
	\begin{center}
		\inputpgf{figs/}{packer-examples.pgf}
		\caption{PACKER generation of a category `B' example, following exposure to one member of category `A' and one member of category `B'. Predictions are shown for three different parameterizations (differing only in $\gamma$): {\em (a)} Predictions based on contrast similarity only. {\em (b)} Predictions based on target similarity only.  {\em (c)} Predictions with both constraints considered.}
		\label{fig:packer-examples}
	\end{center}
\end{figure}

The probability that a candidate $y$ will be generated is evaluated using an Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values of $a(y,x)$ are more likely to be generated than candidates with smaller values:
% 
\begin{equation}
p(y \mid x) = \dfrac
{ \exp  \left \{ \theta \cdot a \left( y, x \right) \right \} } 
{ \sum_i{ \exp  \left \{ \theta \cdot a \left( y_i, x \right) \right\}  } }
\label{eq:packer-choice}
\end{equation}
% 
where $\theta$ ($\theta \geq 0$) is a free parameter controlling response determinism.

It is worth noting that PACKER is only one possible exemplar-based account of category generation within our proposed framework. That is, PACKER places specific constraints on the possible values of $f(x_j)$, but other exemplar-based category generation models with drastically different behavior can be formalized in this framework by imposing alternative constraints. For example, as will be discussed in more detail below, PACKER is formally equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic} when $\gamma = 1$. Likewise, when $\gamma = 0$, PACKER can represent a contrast-only generation mode, relying exclusively on contrast when generating new categories. Finally, when $f(x_j) = -1$ for all $x_j$ (regardless of class membership), a ``pure-packing'' approach is yielded, generating items in unoccupied areas of the domain. Thus, the proposed framework may be used to describe a wide variety of qualitatively distinct generation strategies.

@@

\subsubsection{Relation Between PACKER and Copy-And-Tweak}
\label{section:copytweak-packer}
The PACKER model is similar to the copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are exemplar-based, and both models generate new items according to their similarity to known members of the target class. However, PACKER diverges from the copy-and-tweak model by including a contrast mechanism, enabling generation according to dissimilarity to members of opposing categories. By consequence, copy-and-tweak can be realized as a parameterization of the PACKER model that is insensitive to category contrast. Specifically, when $\gamma = 1$ (see Figure \ref{fig:packer-examples}, panel B), $f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not influenced by these items, and is mathematically equivalent to a copy-and-tweak approach.

In this paper, we report simulations using this copy-and-tweak model. This model fits within the exemplar-based category generation framework defined above, under the constraint that $\gamma = 1$, and is a continuous-dimension adaptation of the model tested by \cite{jern2013probabilistic}. By formalizing a model family where PACKER and copy-and-tweak are different parameterizations of models within the same framework, the comparison between PACKER and copy-and-tweak provides a test of the explanatory value of the contrast mechanism: The account provided by copy-and-tweak will only equal that of PACKER if the contrast mechanism does not offer an advantage (i.e., if $\gamma < 1$ significantly improves model fits). Note that the purpose of the article is to explore and formally analyze the role of contrast in category generation and thus, we leave extending PACKER to incorporate distributional factors \citep[as explored by][]{jern2013probabilistic} for future work.

\subsection{Contrast as representativeness: An extension to Jern and Kemp (2013)}

TODO: fill in Bayesian definition of representativeness and its application in JK2013.

@@

Research on the creative generation of novel concepts has focused on the finding that generated categories tend to possess distributional commonalities with known categories. However, a fundamental goal of concept generation is to create something {\em new} (i.e., different from what is already known). The manner in which generated categories differ from known ones is, nonetheless, poorly understood: Existing theories do not make strong predictions about how creatively generated concepts should systematically differ from existing ones. Above, we found encouraging initial support that contrast influences category generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced a novel, exemplar-based model formalizing the roles of similarity and contrast in creative generation.

@@

\section{Model-based Analyses}

Experiments 1 and 2 revealed systematic and strong effects of category contrast on creative generation. In this section, we report the results of simulations with formal models aimed at explaining our observations. Specifically, we present simulations from the PACKER model, as well as a `copy-and-tweak' model (discussed in Section \ref{section:copytweak-packer}), defined as a variant of PACKER with the $\gamma$ parameter constrained to be one. The comparison of these two models serves to highlight the explanatory role of contrast within PACKER's framework: If contrast affords little explanatory advantage, then the two accounts should produce an equally strong account. We also present simulations from an implementation of the hierarchical Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in Appendix \ref{ap:hsampling-definition}. The comparison between the hierarchical Bayesian model and PACKER is meant to emphasize the necessity of contrast and demonstrate that generation cannot be explained entirely through the emulation of distributional structure. Each model has complementary strengths and weaknesses: Whereas PACKER is insensitive to the distributional structure of learned categories (relying only on within- and between-category similarity), the hierarchical Bayesian model generates categories exclusively on the basis of knowledge of how existing classes are distributed.

@@

Table \ref{table:global-model-fits} contains the AIC values of each model's best fit, as well as the corresponding log-likelihood ($L$) and the best-fitting parameter values. These results reveal strong model differentiation: The PACKER model achieved far better fits compared to the copy-and-tweak and hierarchical Bayesian models, and copy-and-tweak performed somewhat better than the hierarchical Bayesian model. While PACKER's advantage may tentatively be attributed to the model's sensitivity to category contrast (this will be explored in detail below), the advantage shown by copy-and-tweak over the hierarchical Bayesian model may be attributed to its exemplar-based representation, as opposed to the prototype-based representation assumed by the hierarchical Bayesian model. As observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated categories we observed were often widely distributed, with no items near the category prototype. This aspect of the data is inconsistent with the multivariate normal distributions (similar to prototypes) used to represent categories in the \cite{jern2013probabilistic} model, but can be easily accounted for using an exemplar-based approach.

>>>>

%!TEX output_directory = latex_out/
\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

@@

% set up notes-- different backgrounds for Nolan and Joe!
\newcommand\nbcnote[1]{\todo[inline, backgroundcolor = yellow]{\textbf{NBC}: #1}}
\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
\newcommand\sxlnote[1]{\todo[inline, backgroundcolor = cyan]{\textbf{SXL}: #1}}

@@

\vfill
Joseph L. Austerweil\textsuperscript{1}
Nolan Conaway\textsuperscript{1}, 
Shi Xian Liew\textsuperscript{1},
and Kenneth J. Kurtz\textsuperscript{2}
\\[\baselineskip]
\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
\\[1in]

@@

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section*{Abstract}
\jlanote{need to update} 
\sxlnote{updated} 
%
The ability to creatively generate new concepts and ideas is among the most
fascinating aspects of human cognition, but we do not have a strong
understanding of the cognitive processes and representations underlying concept
generation. In this paper, we study the generation of new categories using the
computational and behavioral toolkit of traditional artificial category
learning. Previous work in this domain has focused on how the statistical
structure of known categories generalizes to generated categories, overlooking
the degree of contrast between the known and generated categories as a factor.
We report three experiments demonstrating that contrast between what is known
and what is created is of fundamental importance for categorization and the
creative process. We propose two novel approaches to modeling category contrast:
one focused on exemplar dissimilarity and another on the representativeness
heuristic. These new models are compared to two key alternatives. Our
experiments and simulations demonstrate specifically how category contrast
influences creative generation, and how it can interact with other constraints
to produce different types of creations. Our work also serves as a concrete
example of how well-established and highly controlled experimental frameworks
(such as those in the field of category learning) can be used to make progress
in an intriguing yet understudied domain.

@@

\sxlnote{I think we can address Charles' point on real world examples of how
  contrast can be important around here.}
To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Although implicitly assumed in some work, this constraint has been
overlooked in previous research: To our knowledge, there has not been any
systematic investigation addressing how generated categories {\em differ} from
what is already known. Although the idea of category contrast is ubiquitous
throughout the categorization literature, and extends to a variety of other
fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
be ``different'' is vague, as there are many ways it could be different from a
previously observed category.

We examine two different definitions of how categories should differ from one
another: exemplar dissimilarity or representativeness of the alternative
category. To formalize the first technique, we build on the largely successful
exemplar modeling framework that
\citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
we propose a novel exemplar model of category generation, {\em Producing Alike
  and Contrasting Knowledge using Exemplar Representations} (PACKER),
formalizing how new categories should differ from previous categories. It
embodies the exemplar dissimilarity principle by incorporating a factor that
repulses members of opposite categories from one another. To foreshadow, we find
that this notion of contrast predicts human category learning performance, above
and beyond other category learning models in a novel behavioral experiment. This
demonstrates that contrast plays a central role in other important
categorization tasks, which is not specific to category generation.

The second hypothesis for how contrast might affect categorization is contrast
as {\em representativeness}. The representativeness heuristic \citep{kahneman73}
states that people make judgments regarding an outcome based on how
representative it is of the evidence given in the current context. The heuristic
is a powerful theoretical construct that has been used to capture a range of
complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
especially those that deviate from normative theory (as given by a
straightforward application of Bayes' rule). For example, the coinflip sequence
TTHTH is perceived to be more random (due to it being representative of a
coinflip sequence) than TTTTT (despite both having the same probability of
having been generated from a fair coin). Although there is a healthy debate as
to whether perceived randomness {\em really} deviates from normative theory
\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
main explanations of human judgments \citep{reimers18}. To formalize the
hypothesis that people expect a new category to be representative of the
opposite of the current category, or contrast as {\em representativeness}, we
use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
has been used successfully to capture human performance in color categorization
\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
learning \citep{rafferty10}. We do so by extending the state-of-the-art category
generation model \citep{jern2013probabilistic} and find that human category
generation is best captured as generating representative, rather than probable,
samples.

@@

\sxlnote{Added thomas98 description here} It is worth pointing out that the
observation of regularities in distributional properties across categories is
not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
found that when people were asked to generate feature values for exemplars with
a missing feature (specifically circles of different sizes, each with a radial
line of varying orientation), the resulting set of exemplars tended to share the
same feature correlations as exemplars from previously learned categories. [the
transition to the next point isn't great, though]

@@

\jlanote{check to make sure the next paragraph has been updated enough}
\sxlnote{I think it's updated enough! If given a chance I might edit this a
  little to be less harsh on hierarchical bayesian models in general though I
  think, since the representativeness model is one of them?}
Although it is evident that people are {\em capable} of creating new concepts
and categories, it is not entirely clear how new concepts are systematically
made different from what is already known. The hierarchical Bayesian model
developed by \cite{jern2013probabilistic} assumes that differences between
observed and generated categories are only due to random variation. The model
assumes that generated categories are sampled from the same underlying domain
distribution as observed categories, and will thus share a common distributional
structure. The model does not make predictions about the {\em location} of the
category within the domain (the perceptual instantiation of category members).
Under a strict interpretation of their model, given knowledge of a single
category within the domain, the most probable new category to be generated is
located in {\em exactly} the same location and possesses an identical
distributional structure. This is not an issue with their model specifically,
but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
the latent generative process the same underlying distribution generates all of
the categories and thus, any differences between categories are due to {\em
noise} and should not be {\em systematic}. If the goal of creating a new
category given others in a domain is generating from the posterior predictive
distribution then the best a standard hierarchical Bayesian model without a
notion of contrast built into the model can do at capturing contrast is to
assume that the new category is placed uniformly at random over stimulus space.
But, this defeats the purpose of a hierarchy as it is ignored when determining a
new category location!\footnote{It is plausible that some hierarchical Bayesian
model could be created that generates categories different from each other.
However, this model would not be a standard application or extension of most
pre-existing hierarchical Bayesian models. The generative process would need to
include a component that promotes contrast.} Note that this does not mean
location information and contrast cannot be captured using hierarchical Bayesian
models. In fact, our model of contrast as representativeness is a hierarchical
Bayesian model, but with a different goal than predicting the next category.

@@

\sxlnote{Charles suggested we either remove the section below, or move it to the
  discussion. I've tried finding a good place for it in the discussion but it
  doesn't fit very well, and moving it takes away this nice motivation for
  our further investigation into contrast. I think that keeping it here in the
  introduction might work out, provided Charles finds my argument for why the
  regularity assumption is too strong convincing enough. More details a little
  further down.}
\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}

@@

\sxlnote{Added this bit on regularity below - I'm a little undecided as to leave
  it in the main text or place as a gigantic footnote? It is a bit of a
  digression, and it's really just here to address Charles' point.... If we want
  to keep this in the main text, I'll have to amend the following paragraph
  ``Re-analyzing the results...'' to improve the transition.}
%
It is possible, however, that this data can be explained by a process that does
not involve contrast. Specifically, since participants were trained on exemplars
that share the same hue within a category, it is plausible that they generate
exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
probability that a hue distinct from the experimenter-defined categories is
selected, there is a corresponding $.75$ probability (or about 17 out of 22
participants) that the generated category will have no exemplars that share a
hue with any of the experimenter-defined categories. In addition, that there is
a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
category will have all six exemplars that share a hue with any of the
experimenter-defined categories. This prediction of regularity across categories
is much closer to the data than the prediction from a Binomial model and does
not require any mechanism of category contrast.

However, upon analysing the consistency of hues in the generated categories, it
is clear that participants do not generally produce categories with entirely
similar hues across all exemplars. Specifically, only .55 of all generated
categories comprised exemplars with a common hue. Consequently, the assumption
of regularity across categories may be inappropriate for this data set, leaving
the contrast-based explanation as the most plausible for our purposes.

@@

PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
$\theta_t$ and $\theta_c$ are free parameters
($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
between within- and between-category similarity. For example, when
$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
target category and $f(x_j) = -0.5$ for members of other categories; thus, the
model is likely to generate items that are similar to members of the target
category but are not similar to members of other categories. In this way,
$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
produces exclusive consideration of contrast-category members. The combination
of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
breadth of possible approaches; by fitting it to a dataset, one can describe the
relative roles of between-category contrast and within-category similarity in
generation. See Figure \ref{fig:packer-examples} for an illustration of how
these parameters control the relative influence of within-category similarity
and contrast to other categories when generating a new exemplar.

\begin{figure}
	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
      \caption{PACKER generation of a category `B' example, following exposure
        to one member of category `A' and one member of category `B'.
        Predictions are shown for three different parameterizations (differing
        in values of $\theta_t$ and $\theta_c$): {\em (a)}
        Predictions based on contrast similarity only. {\em (b)}
        Predictions based on target similarity only. {\em (c)} Predictions with
        both constraints considered.}
		\label{fig:packer-examples}
	\end{center}
\end{figure}

The probability that a candidate $y$ will be generated is evaluated using an
Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
of $a(y,x)$ are more likely to be generated than candidates with smaller values:
% 
\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
\right\} } }
\label{eq:packer-choice}
\end{equation}
% 
%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
%determinism.

It is worth noting that PACKER is only one possible exemplar-based account of
category generation within our proposed framework. That is, PACKER places
specific constraints on the possible values of $f(x_j)$, but other
exemplar-based category generation models with drastically different behavior
can be formalized in this framework by imposing alternative constraints. For
example, as will be discussed in more detail below, PACKER is formally
equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
contrast-only generation mode, relying exclusively on contrast when generating
new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
membership), a ``pure-packing'' approach is yielded, generating items in
unoccupied areas of the domain. Thus, the proposed framework may be used to
describe a wide variety of qualitatively distinct generation strategies.

@@

\subsubsection{Relation Between PACKER and Copy-And-Tweak}
\label{section:copytweak-packer} The PACKER model is similar to the
copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
exemplar-based, and both models generate new items according to their similarity
to known members of the target class. However, PACKER diverges from the
copy-and-tweak model by including a contrast mechanism, enabling generation
according to dissimilarity to members of opposing categories. By consequence,
copy-and-tweak can be realized as a parameterization of the PACKER model that is
insensitive to category contrast. Specifically, when $\theta_c = 0$ and
$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
influenced by these items, and is mathematically equivalent to a copy-and-tweak
approach.

In this paper, we report simulations using this copy-and-tweak model. This model
fits within the exemplar-based category generation framework defined above, and
is a continuous-dimension adaptation of the model tested by
\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
copy-and-tweak are different parameterizations of models within the same
framework, the comparison between PACKER and copy-and-tweak provides a test of
the explanatory value of the contrast mechanism based on exemplar dissimilarity:
The account provided by copy-and-tweak will only equal that of PACKER if the
contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
significantly improves model fits). Note that the purpose of the article is to
explore and formally analyze the role of contrast in category generation and
thus, we leave extending PACKER to incorporate distributional factors \citep[as
explored by][]{jern2013probabilistic} for future work.

\subsection{Contrast as representativeness: An extension to Jern and Kemp
(2013)}
\label{section:representativeness-definition}

An alternate conceptualization of category contrast is the idea of
representativeness -- exemplars are generated such that they are more
representative of the target category and thus more distinct from and less
similar to other categories. We adopt the definition specified by
\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
\begin{equation}
  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
\label{eq:representativeness}
\end{equation}

In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
distribution on the hypothesis space that excludes $h$. Note that when there are
only two hypotheses (e.g. two categories) involved in the domain, this prior
takes the value of $1$ since there is ever only one alternative hypothesis.

This Bayesian formalization of representativeness makes it straightforward to
implement into the existing hieriarchical Bayesian model of category generation
developed by \cite{jern2013probabilistic}. Specifically, both models assume that
exemplars are sampled from a given category distribution ($h$ in
Equation \ref{eq:representativeness}). Each category distribution is a
multivariate normal parameterized by a location vector $\mu_k$ and covariance
matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
normal-inverse-Wishart (NIW) distribution parameterized by
$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
\begin{align}
  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
                                              \label{eq:mvnsample}\\
  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
\end{align}

In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
number of dimensions or features in the domain. The $\kappa$ and $\nu$
parameters are freely estimated within the constraints $\kappa > d-1$ and
$\nu > 0$.

With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
with:
\begin{equation}
  \mu_k = \dfrac
    {\kappa\mu_{0} + n_k \bar{x}_k}
    {\kappa + n_k}
    \label{eq:category_mus}
\end{equation}
where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
is the observed category mean. Note that if there are no observed exemplars in
the target category (i.e., if the model is generating a completely novel
category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.

The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
by computing the following:
\begin{equation}
  \Sigma_k = [\Sigma_D\nu + C_k +
    \dfrac
    {\kappa n_k}
    {\kappa + n_k}
    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
  ] (\nu + n_k)^{-1}
  \label{eq:Sigma_y}
\end{equation}
where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
covariance matrix from which $\Sigma_k$ samples are obtained. We can
infer $\Sigma_D$ from the observed category covariances $C$ and the prior
covariance $\Sigma_0$. Specifically,
\begin{equation}
    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
\end{equation}

At this point, both the representativeness model and the hierarchical Bayesian
model from \cite{jern2013probabilistic} are largely identical. However, both
models diverge in their computation of exemplar generation probabilities. While
the original hierarchical Bayesian model produces novel exemplars with
probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
representativeness model generates new exemplars with probabilities weighted by
their representativeness $R(y,h)$. In practice, the probability of generating a
particular candidate $y$ is obtained using an Exponentiated
\citet{luce1977choice} choice rule:
\begin{equation}
  p(y) = \dfrac
    {\exp( \theta \cdot R(y,h))}
    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
\label{eq:representative-choice}
\end{equation}
where $\theta$ is a freely estimated response determinism parameter (constrained
such that $\theta \geq 0$).

Despite both the original hierarchical Bayesian model and our representativeness
model sharing identical hierarchical structures, their distinct response
processes can yield very different exemplars. Specifically, the hierarchical
Bayesian model emphasizes the generation of exemplars that maintain
distributional commonalities across categories. In contrast, the
representativeness model focuses on the generation of exemplars that are
representative of the underlying distribution for a target category. Generally,
this representativeness mechanism results in the generation of exemplars that
are less similar (i.e., more different) to contrast categories. We see this
illustrated in Figure \ref{fig:representative-examples}, where the
representativeness model displays a strong preference for generating exemplars
from target category 'B' that are further away from the contrast category 'A'.

\begin{figure}
	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
		\caption{Generation of a category `B' example, following exposure
to one member of category `A' and one member of category `B'. Predictions are
shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
representativeness model.}
		\label{fig:representative-examples}
	\end{center}
\end{figure}

@@

Research on the creative generation of novel concepts has focused on the finding
that generated categories tend to possess distributional commonalities with
known categories. However, a fundamental goal of concept generation is to create
something {\em new} (i.e., different from what is already known). The manner in
which generated categories differ from known ones is, nonetheless, poorly
understood: Existing theories do not make strong predictions about how
creatively generated concepts should systematically differ from existing ones.
Above, we found encouraging initial support that contrast influences category
generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
novel approaches to formalizing the roles of similarity and contrast in creative
generation.

@@

\section{Model-based Analyses}
\sxlnote{For this section I've added the existence of the representativeness
  model but I think discussion of its performance is still lacking compared to PACKER.}
Experiments 1 and 2 revealed systematic and strong effects of category contrast
on creative generation. In this section, we report the results of simulations
with formal models aimed at explaining our observations. Specifically, we
present simulations from the two novel contrast models: PACKER and the
representativeness model, and compare them to two models that do not incorporate
contrast: the copy-and-tweak model (discussed in Section
\ref{section:copytweak-packer}) and an implementation of the hierarchical
Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
Appendix \ref{ap:hsampling-definition}. For our simulations here, the
copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
parameter constrained to be zero. The comparison of this set of models serves to
highlight the explanatory role of contrast in categorization: If contrast
affords little explanatory advantage, then there should be few differences in
performance between PACKER and copy-and-tweak, or between the representativeness
and hierarchical Bayesian model. The comparison between these models can also
emphasize the necessity of contrast and demonstrate that generation cannot be
explained entirely through the emulation of distributional structure. Each model
has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
are insensitive to the distributional structure of learned categories (relying
only on exemplar similarities), the representativeness and hierarchical Bayesian
model generates categories exclusively on the basis of knowledge of how existing
classes are distributed.

@@

Table \ref{table:global-model-fits} contains the AIC values of each model's best
fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
parameter values. These results reveal strong model differentiation: both
contrast models (PACKER and the representativeness model) achieved far better
fits compared to their non-contrast counterparts: copy-and-tweak and the
hierarchical Bayesian model respectively. Interestingly between the contrast
models, the hierarchical model (representativeness) outperformed the
exemplar-based theory (PACKER), whereas between the non-contrast models, the
reverse is observed. Specifically, here the exemplar-based model
(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.
